```yaml

---
title: "R Markdown Final Project"
author: "Rajdai Bhikam"
date: "2025-04-30"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##This assignment outlines the core functions I created for my final project using the `mtcars` dataset. These functions are part of my R package `RBhikam`, designed to help new R users clean, summarize, and visualize data easily.

---
**FUNCTION 1**

**Purpose:**
This function is used to remove missing values and convert column names to lowercase. 

**Input:**
`df`(data frame): The dataset to clean

**Output:**
A clean dataframe with no `NA` values and standardize column names.

**Code:**

```{r}

clean_data <- function(df) {
  df <- na.omit(df)                 
  names(df) <- tolower(names(df)) 
  return(df)                       
}
``` 
**Example:**
```{r}
cleaned <- clean_data(mtcars)

head(cleaned)

```
**FUNCTION 2**
**Purpose:**
This function groups data by categorical variables and then calculates the mean of a selected column
**Input:**
`df`: The dataframe
`group_col`: column by group
`summary_col`: summary of numeric column

**Output:**
A summary of the datframe

**CODE:**

```{r}
summarize_by_group <- function(df, group_col, summary_col) {
  df |>
    dplyr::group_by(.data[[group_col]]) |>
    dplyr::summarise(mean_value = mean(.data[[summary_col]], na.rm = TRUE))
}
```
**Example:**
```{r}
summarize_by_group(mtcars, "cyl", "mpg")
```

**FUNCTION 3**

**Purpose:**
This function creates a bar plot of the grouped data.

**Input:**
`df`: The dataframe with grouped summaries
`x`: column name for x-axis
`y`: column name for y-axis

**Output:**
A ggplot2 chart

**CODE:**

```{r}
library(ggplot2)

plot_summary <- function(df, x, y) {
  ggplot(data = df, aes(x = .data[[x]], y = .data[[y]])) +
    geom_bar(stat = "identity", fill = "steelblue") +
    theme_minimal()
}
```
**Example:**

```{r}
summary_df <- summarize_by_group(mtcars, "cyl", "mpg")
plot_summary(summary_df, "cyl", "mean_value")


```

**Package Structure**

The RBhikam package contains the following components

- `DESCRIPTION` with metadata and dependencies
- `R/` folder with individual `.R` function files
- `data/mtcars.rda` for the sample dataset
- `README.md` for GitHub overview
- `man/` folder with auto-generated documentation
- `tests/` folder (optional, to test functions)

---

**Installation Instructions**

The package was installed and tested using:

```r
devtools::document()  
devtools::install()    
library(RBhikam)       
```

---


**Reflections & Challenges**

- **What worked:** Function creation and documentation using `roxygen2` was straightforward after the first attempt.
- **What I learned:** I learned how R packages are structured, documented, tested, and published on GitHub.
- **Challenges:** My biggest challenge was fixing documentation errors and making sure functions exported properly. I had a hard time getting everything to work because of the multiple errors. 
- **AI Disclosure:** I used ChatGPT to help debug errors, and clarify package structure.All code was written, tested, and verified manually by me, and I used AI tools as a guide, not as a substitute for learning. The final implementation reflects my own understanding and work.
